Transparency of COVID-19-related research - Analysis

### 2022-03-09 manuscript_covid_transparency.qmd
Results are complete

Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives. Less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: %--%). Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P \< 0.001). Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: %--%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).

A total of 51601 articles were published in journals with no impact factor. Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively). Detailed information is illustrated in Table 1.

Figure 1 shows the monthly trend for three transparency indicators. There was an inconsistent trend for the indicators. However, the COI disclosure percentage seemed to be more consistent and showed an increasing trend over time.

>>>>

Regarding data and code availability, less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: 12.76%--13.25%). Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P=0.1287). Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: 2.98%--3.23%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).

A total of 51601 articles were published in journals with no impact factor. Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with lower JIF (P = 0.003, \< 0.001, and = 0.007, respectively). Detailed information is illustrated in Table 1.

| Measurement       | Mean (SD)       |                | Median (IQR)  |               | P-value  |
|-------------------|-----------------|----------------|---------------|---------------|----------|
|                   | With            | Without        | With          | Without       |          |
| COI disclosure    | 4.097 (2.0086)  | 7.442 (2.1377) | 3.709 (2.253) | 8.483 (0.000) | \< 0.001 |
| Funding statement | 4.085 (1.9627)  | 5.670 (2.8597) | 3.706 (2.253) | 4.773 (4.900) | \< 0.001 |
| Registration      | 3.808 (1.9459	) | 4.396 (2.1769) | 2.863 (1.973) | 4.003 (2.774) | \< 0.001 |
| Data sharing      | 4.411 (2.3977)  | 4.251 (2.1221) | 4.057 (2.545) | 3.763 (2.519) | 0.03679  |
| Code sharing      | 5.891 (4.5176)  | 4.239 (2.0812) | 4.262 (2.927) | 3.800 (2.519) | \< 0.001 |

P-value based on the Wilcoxon rank sum test.

COI: Conflict of interest; SD: Standard deviation; IQR: Inter-quartile range.

Figure 1 shows the monthly trend for three transparency indicators. According to the figure, there was an inconsistent trend for the indicators. However, the COI disclosure and data sharing percentages showed an increasing trend over time.

### 2022-03-09 rmarkdown-covid-19-transparent-analysis.Rmd
Results are complete

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/ahmad/Google Drive/Meta-Research in Dentistry/FAIRness of COVID-19 data/data/')
```

@@

The number and percentage of papers that shared data (preliminary):
```{r}
kable(data.frame(number = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 2)
           ))
```

Comparing data sharing rate in two years (preliminary):
```{r}
kable(round(prop.table(table(opendata$is_open_data, opendata$year_firstpub), 2)*100, 1))
```

The number and percentage of papers that shared code (preliminary):
```{r}
kable(data.frame(number = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 2)
           ))
```

Comparing code sharing rate in two years (preliminary):
```{r}
kable(round(prop.table(table(opendata$is_open_code, opendata$year_firstpub), 2)*100, 1))
```

@@

Total number of papers published in journals with no JIF:
```{r}
nrow(opendata) - nrow(db_jif)
```
JIF based on COI disclosure:

```{r}
kable(db_jif %>% 
  group_by(is_coi_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T),3),
            SD = round(sd(X2020.JIF, na.rm = T),4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

```{r}
kable(db_jif %>% 
  group_by(is_fund_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

```{r}
kable(db_jif %>% 
  group_by(is_register_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

## Monthly trend
Now, drawing monthly trend for each year:
```{r}
summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub) %>%
  summarise(counts = n(),
            coi_true = sum(is_coi_pred == TRUE),
            fund_true = sum(is_fund_pred == TRUE),
            reg_true = sum(is_register_pred == TRUE))

summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts)) + 
  geom_col(position = position_dodge()) +
  facet_grid(year_firstpub ~ .) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12)

@@

coi_plot <- coi_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_coi_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Conflict of Interest disclosure")

@@

fund_plot <- fund_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_fund_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Funding statement")

@@

reg_plot <- reg_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_register_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Registration statement")

library(ggpubr)
figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    reg_plot + theme(strip.text.x = element_blank()),
                    ncol = 1, nrow = 3, vjust = 1, 
                    align = "hv", common.legend = T, legend = "right")

# tiff("Figure.tiff", width = 6, height = 7, units = "in", res = 300)
figure
#dev.off()
```

>>>>

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Documents/GitHub/covid-transparency/data')
```

@@

The number and percentage of papers that shared data:
```{r}
kable(data.frame(number = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 2)
           ))
```

Confidence interval for data sharing:
```{r}

kable(round(epi.prev(pos = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
         tested = nrow(opendata),
         se = 0.758,
         sp = 0.986)$ap, 
      2))
```

Comparing data sharing rate in two years:
```{r}
kable(round(prop.table(table(opendata$is_open_data, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_open_data, opendata$year_firstpub)
```

The number and percentage of papers that shared code:
```{r}
kable(data.frame(number = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 2)
           ))
```

Confidence interval for sharing code:
```{r}

kable(round(epi.prev(pos = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
         tested = nrow(opendata),
         se = 0.587,
         sp = 0.997)$ap, 
      2))
```

Comparing sharing code rate in two years:
```{r}
kable(round(prop.table(table(opendata$is_open_code, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_open_code, opendata$year_firstpub)
```

@@

Total number of papers published in journals with no JIF:
```{r}
nrow(opendata) - nrow(db_jif)
```

JIF based on COI disclosure:
```{r}
kable(db_jif %>% 
  group_by(is_coi_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T),3),
            SD = round(sd(X2020.JIF, na.rm = T),4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

JIF based on funding disclosure:
```{r}
kable(db_jif %>% 
  group_by(is_fund_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

JIF based on registration:
```{r}
kable(db_jif %>% 
  group_by(is_register_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

@@

JIF based on data sharing:
```{r}
kable(db_jif %>% 
  group_by(is_open_data) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

with(db_jif, shapiro.test(X2020.JIF[is_open_data == TRUE]))
with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == FALSE], 5000)))
```

Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(X2020.JIF~is_open_data, data = db_jif, exact = FALSE)
```

JIF based on code sharing:
```{r}
kable(db_jif %>% 
  group_by(is_open_code) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

with(db_jif, shapiro.test(X2020.JIF[is_open_code == TRUE]))
with(db_jif, shapiro.test(sample(X2020.JIF[is_open_code == FALSE], 5000)))
```

Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(X2020.JIF~is_open_code, data = db_jif, exact = FALSE)
```

## Monthly trend
Now, drawing monthly trend for each year:
```{r}
summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub) %>%
  summarise(counts = n(),
            coi_true = sum(is_coi_pred == TRUE),
            fund_true = sum(is_fund_pred == TRUE),
            reg_true = sum(is_register_pred == TRUE),
            data_true = sum(is_open_data == TRUE),
            code_true = sum(is_open_code == TRUE))

@@

coi_plot <- coi_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_coi_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Conflict of Interest disclosure",
       y = "Number of articles") +
        scale_fill_discrete(labels = c("Without", "With"),
                            guide = guide_legend(reverse=TRUE))

@@

fund_plot <- fund_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_fund_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Funding statement",
       y = "Number of articles") +
        scale_fill_discrete(labels = c("Without", "With"),
                            guide = guide_legend(reverse=TRUE))

@@

reg_plot <- reg_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_register_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Registration statement",
       y = "Number of articles") +
        scale_fill_discrete(labels = c("Without", "With"),
                            guide = guide_legend(reverse=TRUE))

data_summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub, is_open_data) %>%
  summarise(counts = n())

data_plot <- data_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_data)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Data sharing",
       y = "Number of articles") +
        scale_fill_discrete(labels = c("Without", "With"),
                            guide = guide_legend(reverse=TRUE))

code_summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub, is_open_code) %>%
  summarise(counts = n())

code_plot <- code_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_code)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Code sharing",
       y = "Number of articles") +
        scale_fill_discrete(labels = c("Without", "With"),
                            guide = guide_legend(reverse=TRUE))

library(ggpubr)

figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    reg_plot + theme(strip.text.x = element_blank()),
                    ncol = 1, nrow = 3, vjust = 1, 
                    align = "hv", common.legend = T, legend = "right")

figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    reg_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    data_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    code_plot + theme(strip.text.x = element_blank()),
                    ncol = 1, nrow = 5, vjust = 1, 
                    align = "hv", common.legend = T, legend = "right")

# tiff("Figure.tiff", width = 6, height = 11, units = "in", res = 300)
figure
#dev.off()
```
